---
title: "💻 **Introduction to VIMuRe**"
subtitle: "VIMuRe v0.1.0 (latest)."
categories: [basics]
tags: 
  - basics
  - tutorial
  - R
description: "Read to understand the data input format expected by the <code>vimure</code>, and how to run and  interpret the output of the model."
---

<div style="display:inline-block;border-radius: 0.25rem;border: solid 1px #dee2e6;margin:1%;padding:1.5%;background-color:$popover-bg">

If you use `vimure` in your research, please cite the following paper:

> De Bacco, Caterina, Martina Contisciani, Jonathan Cardoso-Silva, Hadiseh Safdari, Diego Baptista, Gabriela L. Borges, Tracy Sweet, et al. 'Latent Network Models to Account for Noisy, Multiply-Reported Social Network Data'. arXiv, 12 December 2022. doi:[10.48550/arXiv.2112.11396](https://arxiv.org/abs/2112.11396). [@de_bacco_latent_2022]

</div>


(🚧 This entire tutorial is a work in progress. 🚧)

# ⚙️ Setup

**Import packages**

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(knitr)
```

```{r setup, warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)

library(igraph)
library(vimure)
```

# 📥 Load the data

Here, we'll use the `vimure` package to model social networks from a village in Karnataka, India, gathered as part of a project studying the adoption of microfinance [@banerjee_diffusion_2013]. The data files we use here are derived from those openly-available on the Harvard Dataverse page [@banerjee_diffusion_2013-1].

We have selected a particular village to focus on. The dataset contains information on multiple different types of relationships. We are interested in relationships that were "double-sampled" in the original survey, meaning those elicited through two different prompts. For example, respondents were asked both _"If you needed to borrow kerosene or rice, to whom would you go?"_ as well as _"Who would come to you if he/she needed to borrow kerosene or rice?"_ These two different prompts should give us different perspectives on what could be the same relationship (i.e., we're asking $i$ if they **go** to $j$ for help, and also asking $j$ if $i$ **comes** to them for help). 

Four different tie types (or "layers") were double-sampled in this way (i.e., elicited with two prompts). Here, we will focus just on one tie type: the "visiting" relationship (based on two prompts: _"In your free time, whose house do you visit?" and "Who visits your house in his or her free time?"_). 

::: panel-tabset

## Load data.frame

After our preprocessing, the input data takes the form of the dataset given below. 

- `ego`: generally the requester, also referred to as $i$ in [@de_bacco_latent_2022] (required ✔️)
- `alter`: generally the giver, also referred to as $j$ in [@de_bacco_latent_2022] (required ✔️)
- `reporter`: the person reporting on the tie, as referred to as $m$ in [@de_bacco_latent_2022] (required ✔️)
- `prompt_type`: the prompt type that was given (optional)
- `layer`: the tie type (optional)
- `weight`: the weight of the edge (optional)

Suppose we have a data frame `df_edges` of our village's data that looks like this:

```r
head(df_edges, 10)
```

```{r, echo = FALSE}
URL <- "https://raw.githubusercontent.com/latentnetworks/vimure/develop/data/input/india_microfinance/formatted/"
village <- "30" # change this parameter to run in another village

df_metadata <- read.csv(paste0(URL, paste0("vil", village, "_meta.csv")))
df_edges <- read.csv(paste0(URL, paste0("vil", village, "_edges.csv")), stringsAsFactors = F)

df_edges$layer <- factor(df_edges$type, 
  levels=c(
    "borrowmoney","receivemoney", "receiveadvice", "helpdecision",
    "keroricego", "keroricecome", "visitgo", "visitcome"
    ), labels=c("money", "money", "advice", "advice", "kerorice", "kerorice", "visit", "visit"))
layer <- "visit" # Change this to run for a different layer
df_edges <- df_edges[df_edges$layer %in% layer, ]
df_edges <- df_edges[df_edges$i != df_edges$j, ]

df_edges <- 
  df_edges %>% 
  rename(m=respondent, prompt_type=`type`) %>% 
  select(i, j, m, prompt_type, layer, weight) 
  
df_edges$prompt_type <- factor(df_edges$prompt_type, levels=c("visitgo", "visitcome"))

# Only keep reports made by those who were MARKED as respondents in metadata CSV
respondents <- df_metadata %>% filter(didsurv == 1) %>% pull(pid)
df_edges <- df_edges %>% filter(i %in% respondents & j %in% respondents) %>% ungroup()

df_edges <- df_edges %>% rename(ego=i, alter=j, reporter=m)

df_edges %>% 
  head(10) %>% 
  knitr::kable(row.names=FALSE, align = "cccccc", caption = "A typical data input for VIMuRe", full_width = FALSE)

```

## Load `igraph` object

Alternatively, you might have your data as a **directed** `igraph` object. In this case, you must ensure that the `igraph` object has the following attributes:

- `reporter`: the person reporting on the tie, as referred to as $m$ in [@de_bacco_latent_2022] (required ✔️)
- `prompt_type`: the prompt type that was given (optional)
- `layer`: the tie type (optional)
- `weight`: the weight of the edge (optional)

Note that the `graph_from_data_frame` function assumes that the first two columns in the data frame comprise the edgelist (i.e., the $ego$ and $alter$ columns and treats all subsequent columns as edge attributes). 

```{r, echo = FALSE}
G <- igraph::graph_from_data_frame(df_edges, directed=TRUE)
```

Suppose our village's data is in the form of an `igraph` object `G`:

```{r}
G
```

:::

## 📊 Summary statistics

### Number of nodes

::: panel-tabset

## data.frame

```{r}
dplyr::union(df_edges$ego, df_edges$alter) %>% n_distinct()
```

## `igraph` object

```{r}
igraph::vcount(G)
```

:::


### Number of reporters

::: panel-tabset

## data.frame

```{r}
df_edges$reporter %>% n_distinct()
```

## `igraph` object

```{r}
igraph::get.edge.attribute(G, "reporter") %>% unique() %>% length()
```

:::

### Average number of ties per reporter

There are a couple of things to note about this dataset: the maximum number of name generators is four, and reporters can only report on ties that they are a part of (e.g., if a reporter is not a part of a tie, they cannot report on it). Because we are modeling double-sampled questions, each reporter can report a maximum of 8 ties.


::: panel-tabset

## data.frame

```{r, fig.width=8, fig.height=4}
plot_df <- 
  df_edges %>% 
  group_by(prompt_type, reporter) %>% 
  summarise(n = n(), .groups="keep")

# Sort n in descending order
plot_df$n <- factor(plot_df$n, levels=sort(unique(plot_df$n), decreasing=TRUE))


g <- 
(
  ggplot(plot_df, aes(x = n, fill=prompt_type)) +
  geom_bar() +

  scale_fill_manual(values = c("visitgo" = "#1f77b4", "visitcome" = "#ff7f0e"),
                    labels = c("visitgo" = "Whose house do you visit?", 
                               "visitcome" = "Who visits your house?")) +
  coord_flip() +

  theme_bw() +
  theme(legend.position = "bottom") +
  labs(x = "Number of ties per reporter", 
       y = "Number of reporters",
       fill = "Prompt type")
)
g
```

## `igraph` object

```{r, fig.width=8, fig.height=4}
plot_df <- 
  bind_cols(reporter=E(G)$reporter, prompt_type=E(G)$prompt_type) %>%
  group_by(prompt_type, reporter) %>% 
  summarise(n = n(), .groups="keep")

# Sort n in descending order
plot_df$n <- factor(plot_df$n, levels=sort(unique(plot_df$n), decreasing=TRUE))

g <- 
(
  ggplot(plot_df, aes(x = n, fill=prompt_type)) +
  geom_bar() +

  scale_fill_manual(values = c("visitgo" = "#1f77b4", "visitcome" = "#ff7f0e"),
                    labels = c("visitgo" = "Whose house do you visit?", 
                               "visitcome" = "Who visits your house?")) +
  coord_flip() +

  theme_bw() +
  theme(legend.position = "bottom") +
  labs(x = "Number of ties per reporter", 
       y = "Number of reporters",
       fill = "Prompt type")
)
g
```

:::

### Union vs. Intersection and the concept of Concordance

Concordance is the proportion of the ties in the network that both reporters report. It measures the extent to which the two reporters agree about the network [@ready_measuring_2021]. It is calculated as follows:


$$
\text{Concordance} = \frac{\text{\# of ties reported by both reporters}}{\text{\# number of unique ties reported}} 
$$


::: panel-tabset

## data.frame

```{r, message=FALSE}
# Take the intersection: keep only records where 
# both reporters report on the same tie in both prompt_types
df_intersection <- 
  df_edges %>% 
  group_by(ego, alter, .drop=FALSE) %>% 
  filter(n() == 2) %>% 
  select(ego, alter) %>%
  distinct()

# Take the union: keep all ties reported 
# irrespective of prompt_type and how many times they were reported
df_union <- df_edges %>% select(ego, alter) %>% distinct()

# Concordance
cat(paste0("Concordance is: ", nrow(df_intersection) / nrow(df_union)))
```


## `igraph` object

```{r}
# Take the intersection: keep only records where
# both reporters report on the same tie in both prompt_types
which_edges_intersection <- E(G)[count_multiple(G, E(G)) == 2]

G_intersection <- 
  igraph::subgraph.edges(G, which_edges_intersection) %>%
  igraph::simplify()

# Take the union: keep all ties reported
# irrespective of prompt_type and how many times they were reported

G_union <- igraph::simplify(G)

# Concordance
cat(
  paste0(
    "Concordance is: ", 
    igraph::ecount(G_intersection) / igraph::ecount(G_union)
  )
)
```
:::


# 📦 Use `vimure` package to run the model

::: panel-tabset

## data.frame

**Convert the data.frame to a `vimure` graph object**

When working directly with data frames representing an edgelist, we first need to use the `parse_graph_from_edgelist()` function to parse the edgelist into a `vimure` object. It is a good practice to specify the name of the columns representing the ego, alter, and reporter, especially if the column names are not the default ones expected by the package.


```{r, message=FALSE}
library(vimure)

vm_graph <- parse_graph_from_edgelist(df_edges, ego="ego", alter="alter", reporter="reporter")

vm_graph
```

::: {.callout-important}

Note that `M = 159` contrasts with the `M = 150` we saw in the [Summary Statistics section](#summary-statistics) above. This is because, by default, `vimure` assumes that all nodes in the network were interviewed and that they _could_ report on each ego-alter pair, even if they have not reported any ties.

If we had a different survey setup, we would need to specify the `R` parameter in the `parse_graph_from_edgelist()` function. We are adding a tutorial in the future to discuss how to handle other survey setups.
:::

**Now we can run the model**

```{r, message=FALSE}
#TODO: refactor `vimure` function to accept a `BaseNetwork` object directly 
# (no need to pass the `x` attribute here)
model <- vimure(vm_graph$X, num_realisations=5, max_iter=100, seed=1, mutuality = TRUE)
```

## `igraph` object

As long as your graph `G` has the edge attributes `E(G)$reporter` and `E(G)$prompt_type`, you can run the model directly on the `igraph` object:

```{r, message=FALSE}
model <- vimure(G, num_realisations=5, max_iter=100, seed=1, mutuality = TRUE)
```

:::

The parameters `num_realisations` and `max_iter` control the number of iterations of the Variational Inference algorithm. The `seed` parameter sets the seed for the random number generator. The `mutuality` parameter specifies whether the model should assume that ties are mutual, as described in [@de_bacco_latent_2022].

## Model Summary

Various elements from the VIMuRe model's output may be of interest:

```{r}
model_summary <- summary(model)
model_summary
```

The summary above shows the model parameters' priors, posteriors, and estimates. For now, let's just focus on a few elements that result from this model summary:

```{r}
names(model_summary)
```

To better understand this summary, refer to the [Advanced: Optimisation Diagnostics](#advanced-optimisation-diagnostic) section below.

## Estimated reliability

Let's first consider the resulting estimates of the "reliability" of each reporter from a data frame available at `model_summary$reliability`. Reliability is represented as $\theta$ in the model [@de_bacco_latent_2022] and is the probability that a reporter will report a tie between two nodes. The model estimates the reliability values and are not directly observed in the data.

While interpretation isn't always straightforward, we can get a general sense of the extent to which particular reporters seem to be "over-" or "under-" reporting their ties. Roughly speaking, nodes with reliability values close to 0 are likely to be under-reporting, while nodes with large reliability values are likely to be over-reporting. 

```{r fig.height=3.75, fig.width=10}
#TODO: Rename `layer` to `theta_value` in the reliability data frame
plot_df <- model_summary$reliability %>% rename(theta_value = layer)

ggplot(plot_df, aes(x = reorder(factor(node), -theta_value), y=theta_value)) +
  geom_point(alpha = 0.4, size=4, color = "#c45241") +
  theme_bw() +
  labs(x = "Nodes sorted from over-to under- reporting", 
       y = "Realibility values") +
  theme(axis.text.x=element_blank())
```

## Mutuality

Another central concept of our model is mutuality, the tendency to report ties in both directions involving the same alter, represented by the parameter $\eta$ in the model. In the paper [@de_bacco_latent_2022], we argue that mutuality is a fundamental property of social ties and that it is important to model it in a latent network model.  We can extract the geometric expectation of $\eta$ from the fitted model object as follows:

```{r}
model$G_exp_nu
```

A mutuality of $\eta \approx 0.78$ is high, which means the model estimates that the average person is _very_ likely to nominate the same alter when prompted in both directions of the tie. They will likely answer the same set of people to the following questions: _"Whose house do you visit?"_ as well as _"Who visits your house?"_. 


# 🕸️ Extracting an inferred network

## Point estimates

After running the model, we can extract the point estimates of the network from the model. This can be done by parsing the $\rho$ values inferred by the model, which can be interpreted as the probability that a tie exists between two nodes. 

If you want to look at $\rho$ directly, you can invoke `model$rho` but notice you will still have to parse the $\rho$ values to get a network object. This is because the $\rho$ values are stored in a 4-dimensional array, where the first dimension is the layer, the second and third dimensions are the nodes ($i\rightarrow j$), and the fourth dimension reflects the expected degree of that tie (up to degree $K$):

```{r}
model$rho %>% dim()
```

In other words, $\rho_{lijk}$ is the probability that a tie degree $k-1$ exists between nodes $i$ and $j$ in layer $l$.

(The last dimension is also related to the parameter $\lambda_k$. You can read more about it on the paper [@de_bacco_latent_2022] as well as on the [Advanced: Optimisation Diagnostics](#advanced-optimisation-diagnostic) section below).

Because it might be tricky to interpret the $\rho$ array directly, we recommend using the `get_inferred_model()` function to parse the $\rho$ values into an adjacency matrix. This function allows several ways to obtain the inferred network, and its usage depends on the interpretation you want to give to the $\rho$ values.

## Sampling from the posterior

(🚧 This section is a work in progress. 🚧)

# 📈 Advanced: Optimisation Diagnostic

(🚧 This section is a work in progress. 🚧)

---------------------------