---
title: "Karnataka Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Karnataka Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vimure)
library(igraph)
library(reshape2)
library(ggplot2)
```

The goal of this article is to run the `vimure` model on the real dataset Karnataka.

# Read the edgelist

First you'll need to read a edgelist.

```{r}
URL <- "https://raw.githubusercontent.com/latentnetworks/vimure/develop/data/input/india_microfinance/formatted/"
village <- "12" # change this parameter to run in other village

PATH <- paste0(URL, paste0("vil", village, "_edges.csv"))
df_edges <- read.csv(PATH, stringsAsFactors = F)
head(df_edges)
```
Then, run the code below to preprocess the data, removing self-loops and re-labeling layers names.

```{r}
df_edges$layer <- factor(df_edges$type, 
  levels=c(
    "borrowmoney","receivemoney", "receiveadvice", "helpdecision",
    "keroricego", "keroricecome", "visitgo", "visitcome"
    ), labels=c("money", "money", "advice", "advice", "kerorice", "kerorice", "visit", "visit"))

layer <- "visit" # Change this to run for a different layer
df_edges <- df_edges[df_edges$layer %in% layer, ]
df_edges <- df_edges[df_edges$i != df_edges$j, ]
head(df_edges)
```

## Using an igraph network

Vimure models expects as input an igraph object that can contains the _layer_, _reporter_ and _weight_ of each edge. 
Use `igraph::graph_from_data_frame` for create a graph using a dataset. The first two columns are the edgelist. Additional columns are considered as edge attributes (See more in ?graph_from_data_frame).

> Vimure extracts these attributes by name, so it is crucial to rename the attributes before creating the graph (or parse the edgelist directly by `parse_graph_from_edgelist` and use params ego,alter,.. for mapping columns names - section below):

```{r}
edges <- df_edges[c(2:5, 1)] # Reorder columns
names(edges) <- c("Ego", "Alter", "layer", "weight", "reporter") # Rename
graph <- igraph::graph_from_data_frame(edges)

graph <- set_graph_attr(graph, "layout", layout_with_fr(graph, niter=600))
plot(graph,
     vertex.size=2,
     vertex.label=NA,
     edge.width=0.5,
     edge.arrow.size=0.5, 
     edge.arrow.width=0.5)
```

Calculate some network statistics from the adjacency matrix:

```{r}
message(
  "Nodes: ", length(V(graph)),
  " | Edges: ", gsize(graph),
  " | Avg. degree: ", mean(degree(graph)), #directed graph,
  " | Reciprocity: ", reciprocity(graph)
)
```

Run the model:

```{r}
# Graph input
model <- vimure(graph, num_realisations=1L, max_iter=100L, seed=5, mutuality = T)
```


## Parsing data to vimure classes
 
Vimure models expect as input the attributes `X` and `R` of a `parse_graph_from_edgelist` result. 
`parse_graph_from_edgelist` expects a dataframe with two required columns (_Ego_ and _Alter_) and three optional columns ( _layer_, _reporter_ and _weight_). Rename the dataframe's columns or use params ego,alter,... of `parse_graph_from_edgelist` for mapping column names to convert.

```{r}
vm_graph <- parse_graph_from_edgelist(df_edges, ego="i", alter="j", reporter="respondent")
```

And run the model:
```{r}
model <- vimure(vm_graph$X, vm_graph$R, num_realisations=1L, max_iter=100L, seed=5, mutuality = T)
```

# Diagnostics

```{r}
diag <- summary(model)
```

## Elbo values

```{r}
df_elbo <- diag$trace

ggplot2::ggplot(df_elbo, ggplot2::aes(x=iter, y=elbo, group=realisation, color=factor(realisation))) +
  ggplot2::geom_point() +
  ggplot2::geom_line(size=1.2) +
  ggplot2::theme_bw() +
  ggplot2::labs(x = "Iter", y="ELBO", color="Realisation")
```

## Estimated reliability

```{r}
plot_df <- reshape2::melt(diag$reliability, id = "node")

ggplot2::ggplot(plot_df, aes(x = reorder(factor(node), -value), y=value)) +
  ggplot2::geom_point(alpha = 0.4, size=2, color = "#c45241") +
  ggplot2::theme_bw() +
  ggplot2::labs(x = "Nodes sorted from over-to under- reporting", y = "Realibility values") +
  ggplot2::theme(axis.text.x=element_blank())
```
## Reconstruct Y matrix 

```{r}
Y_hat <- get_inferred_model(model)

graph <- graph_from_adjacency_matrix(Y_hat[1, , ], mode = "directed") 
graph <- set_graph_attr(graph, "layout", layout_with_fr(graph, niter=600))
plot(graph,
     vertex.size=2,
     vertex.label=NA,
     edge.width=0.5,
     edge.arrow.size=0.5, 
     edge.arrow.width=0.5)
```

