---
title: "Introduction to vimure"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to vimure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of vimure is to make available latent network models to account for noisy, multiply-reported social network Data.

```{r setup}
library(vimure)
library(igraph)
library(ggcorrplot)
```

# Generate synthetic data

First create an object with the desired synthetic network class:

```{r}
net <- gm_Multitensor(N=100, M=50, L=1, C=2, K=2, avg_degree=10, seed=24, sparsify=T, eta=0.2)

# Ground Truth Network
Y <- extract_Y(net)
layer <- 1 # Change to plot others layers for multiply-layers networks

graph <- graph_from_adjacency_matrix(Y[layer, , ], mode = "directed") 
graph <- set_graph_attr(graph, "layout", layout_with_fr(graph, niter=600))
plot(graph,
     vertex.size=5,
     vertex.label=NA,
     edge.width=0.5,
     edge.arrow.size=0.5, 
     edge.arrow.width=0.5)
```

## Python interface

`vimure` is a R binding of a Python package. Many Python basic objects are quickly converted to R automatically.
Custom Python objects that can not be converted automatically are stored in R as a `python.builtin.object`.
As a `python.builtin.object`, you can access all object's attributes as it is in Python using the dollar sign `$`.

Use the function `class` to check if a object is stored in Python.

```{r}
class(net)
```

`net` is stored as a Python object. You can access its attributes using the dollar sign `$` or using our `extract_*` functions which always will return a R object.

```{r}
net$Y # still a python object because it is a sptensor
Y <- extract_Y(net) # extract_Y convert to array
```

Calculate some network statistics from the adjacency matrix:

```{r}
message(
  "Nodes: ", length(V(graph)),
  " | Edges: ", gsize(graph),
  " | Avg. degree: ", mean(degree(graph)), #directed graph,
  " | Reciprocity: ", reciprocity(graph)
)
```

Then, given a network Y, we can generate N observed adjacency matrices as would have been reported by reporting nodes $m$ for $m \in N$ This is achieved by the function `build_X`. Example:

```{r}
X <- build_X(net, flag_self_reporter=T, cutoff_X=F, seed=10L)
dim(X)
```
Use the code below to explore the network from the point of view of different reporting nodes:

```{r}
reporter_id <- 31
message("Reporter ID=", reporter_id," | theta= ", net$theta[layer, reporter_id])

X_current <- X[layer,,,reporter_id]
X_current <- X_current[rowSums(X_current[,-1]) != 0, colSums(X_current[-1,]) != 0]

ggcorrplot( X_current) + 
   ggplot2::scale_fill_gradient(low="white",high="#003396")
```

# Run model

In R we can construct and fit the vimure model by using the `vimure` function.
The `vimure` function inherit all arguments from the original `VimudeModel` class and `VimureModel.fit()`.
See more info about arguments in `help(vimure)`.

In R, the `vimure` takes as data input an adjancency matrix or an igraph object.

```{r}
model <- vimure(X, net$R, mutuality=F, K=2, num_realisations=1, max_iter=150)
```

> Note: Run model without mutuality is not particularly useful since we know this network has some level of reciprocity and mutuality. We are doing this for testing purposes only.

## Diagnostics

Use `summary` to produce result summaries of the results of the model. The output of `summary` is a R list that contains vimure's priors and posteriors and more. You can access the results to plot and analyse the final model. 
Open an issue in our [repository](https://github.com/latentnetworks/vimure/issues/new) to make request new metrics in `summary`.

```{r}
diag <- summary(model)
```

