---
title: "Synthetic data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synthetic data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vimure)
library(ggplot2)
library(ggcorrplot)
library(igraph) 
```

The goal of this article is to illustrate how to generate and use synthetic networks.

# Generating Ground Truth Networks - `Y`

Simply create an object with the desired synthetic network class (more info about arguments and others synthetic models in `help(gm_Multitensor)`):

```{r}
random_net <- gm_Multitensor(N=100, M=100, L=1, C=2, avg_degree=10, sparsify=T, eta=0.99, seed=10)

Y <- extract_Y(random_net)

ggcorrplot(Y[1, ,]) + 
   ggplot2::scale_fill_gradient(low="white",high="#003396")
```

Plot the network data:

```{r}
graph <- graph_from_adjacency_matrix(Y[1, , ], mode = "directed")
graph <- set_graph_attr(graph, "layout", layout_with_fr(graph, niter=600))
plot(graph,
     vertex.size=5,
     vertex.label=NA,
     edge.width=0.5,
     edge.arrow.size=0.5, 
     edge.arrow.width=0.5)
```

Calculate some network statistics from the adjacency matrix:

```{r}
message(
  "Nodes: ", length(V(graph)),
  " | Edges: ", gsize(graph),
  " | Avg. degree: ", mean(degree(graph)), #directed graph,
  " | Reciprocity: ", reciprocity(graph)
)
```

# Generating Observed Networks - X

Given a network Y, we can generate N observed adjacency matrices as would have been reported by reporting nodes $m$ for $m \in N$. This is achieved by the function `build_X`. Example:

```{r}
X <- build_X(random_net, flag_self_reporter=T, cutoff_X=F, seed=10L)
dim(X)
```

## 'Reliability' parameter $\theta$ - drawn from a gamma distribution

> Nodes are not reliable and will report imperfect views of the ground truth network. 

The 'reliability' of nodes is represented by the parameter $\theta_{lm}$ and by default are modeled as a gamma function with shape `sh_theta` and scale `sc_theta`:

$$
\theta_{lm} \sim \Gamma(sh\_theta, sc\_theta)
$$

The function `build_X` uses `sh_theta = 2.0` and `sc_theta = 0.5` by default.

```R
build_X(random_net, sh_theta=2.0, sc_theta=0.5) # Default values
```

The distribution looks as follow:

```{r}
theta <- rgamma(2000, shape = 2, scale = 0.5)
hist(theta, breaks=50, prob = T, xlab = "Reliability parameter (theta)")
lines(density(theta), col="red", lwd=2) 
```

> This setup aims to represent 'reliable' reporters as having $\theta = 1.0$. Reporters with very low $\theta$ ($\theta \to 0$) will tend to under-report while those with high $\theta$ (unbounded) will over-report their social ties.

```{r}
layer <- 1
reporter_id <- 31 # Change this to view the network from the point of view of different reporters

message("Reporter ID=", reporter_id," | theta= ", random_net$theta[layer, reporter_id])

X_current <- X[layer,,,reporter_id]
X_current <- X_current[rowSums(X_current[,-1]) != 0, colSums(X_current[-1,]) != 0]

ggcorrplot( X_current) + 
   ggplot2::scale_fill_gradient(low="white",high="#003396")
```

If we take the average of X (considering the reporter's mask), it will look like this:

```{r}
Xavg <- extract_Xavg(random_net)

ggcorrplot( Xavg[layer, ,]) + 
   ggplot2::scale_fill_gradient(low="white", high="#003396")
```

## 'Reliability' parameter $\theta$ - custom

Instead of drawing $\theta$ from a Gamma distribution, one could just simply pass a list of reliability parameters per each node. It is used to simulate exaggeration scenarios (under-reporters and over-reporters).

```{r}
theta_ratio <- 0.20 # percentage of reporters that are exaggerators
exaggeration_type <- "over" # exaggerators are over-reporters

custom_theta <- build_custom_theta(
    random_net,
    theta_ratio=theta_ratio,
    exaggeration_type=exaggeration_type,
    seed = 1L
)

custom_theta
```

If using a custom theta, pass it to function build_X:

```{r}
X_custom <- build_X(
  random_net,
  flag_self_reporter=T,
  seed=10L,
  theta=custom_theta, 
  cutoff_X=T
)

Xavg <- extract_Xavg(random_net)
ggcorrplot( Xavg[1, ,]) + 
   ggplot2::scale_fill_gradient(low="white", high="#003396")
```

## Mutuality Parameter $\eta$

In our model, we simulate double-sampled reports. The same reporter $m$ responds about giving and receiving social support about two nodes $i$ and $j$. Therefore, every tie type is sampled twice for each reporter, one for each direction of the interaction: $i \to j$ and $j \to i$.

The mutuality parameter ($\eta$) models people’s propensity of naming of the same individuals across both prompts.

**An example of impossibly High Mutuality on X**

```{r}
X_custom <- build_X(random_net, cutoff_X=T, mutuality=0.99, flag_self_reporter=T, seed=10L)
Xavg <- extract_Xavg(random_net)
ggcorrplot( Xavg[1, ,]) + 
   ggplot2::scale_fill_gradient(low="white", high="#003396")
```

**An example of NO mutuality on X ($\eta = 0.0$)**

```{r}
X_custom <- build_X(random_net, cutoff_X=T, mutuality=0, flag_self_reporter=T, seed=10L)
Xavg <- extract_Xavg(random_net)

ggcorrplot( Xavg[1, ,]) + 
   ggplot2::scale_fill_gradient(low="white", high="#003396")
```

**Remember:** These X adjacency matrices have all been generated for the exact same ground truth Y. One could play around with the parameters a bit further to investigate or imagine different scenarios of reporting inaccuracy.
