---
title: "RMarkdown 2"
author: "Gabriela L. Borges"
date: "5/18/2022"
output: html_document
---

**OBJECTIVE:** Illustrate how to run the foundational model with `vimure` function

```{r setup, include=FALSE}
library(reshape2)
library(vimure)
library(ggplot2)
library(ggcorrplot)
library(igraph) 
library(devtools)
```

# Generate network

## Generate Y

(see RMarkdown 01 for more information about the synthetic network models used in the study)

Let us generate a synthetic network with reciprocity (`gm_Multitensor`):

```{r}
# Eta is the expected reciprocity on the network
random_net <- gm_Multitensor(N=100, M=100, L=1, C=2, K=2, avg_degree=10, seed=24, sparsify=T, eta=0.2) 

layer <- 1
Y <- extract_Y(random_net)

ggcorrplot(Y[layer, ,]) + 
   scale_fill_gradient(low="white",high="#003396")
```

What is the actual overall reciprocity of the network?

```{r}
overall_reciprocity(Y[layer, , ])
```
## Generate X

```{r}
X <- build_X(random_net, cutoff_X=T, mutuality=0.3, lambda_diff=1.0, flag_self_reporter=T)
Xavg <- extract_Xavg(random_net)

ggcorrplot(Xavg[layer, ,]) + 
   scale_fill_gradient(low="white",high="#003396")
```

## Distribution of $\theta$

```{r}
theta_df <- data.frame(layer=t(random_net$theta))
theta_df <- melt(theta_df)

ggplot(theta_df, aes(x = value)) +
  geom_histogram() +
  xlab("theta") +
  ylab("# of nodes") +
  facet_grid(~variable) +
  theme_bw() +
  ggtitle(
    paste0("Distribution of theta accross different layers of the generated network. (N=",
           random_net$N,", M=", random_net$M,", C=", random_net$C,")"))

```

# Baseline

## Union of X

```{r}
X_union <- random_net$X_union$toarray()
ggcorrplot(X_union[layer, ,]) + 
   scale_fill_gradient(low="white",high="#003396")
```

## Intersection of X

```{r}
X_intersection <- random_net$X_intersection$toarray()
ggcorrplot(X_intersection[layer, ,]) + 
   scale_fill_gradient(low="white",high="#003396")
```
# VIMuRe model

In R we can construct and fit the vimure model by using the `vimure` function.
The `vimure` function inherit all arguments from the original `VimudeModel` class and `VimureModel.fit()`.
See more info about arguments in `help(vimure)`.

In R, the `vimure` takes as data input an adjancency matrix or an igraph object.

## Run model without mutuality ($\eta = 0$)

This is not particularly useful since we know this network has some level of reciprocity and mutuality.

### Using Igraph

```{r}
graph <- graph_from_adjacency_matrix(Xavg[1,,])
model <- vimure(graph, mutuality=F, K=2, num_realisations=1, max_iter=150)
```


### Using X matrix

```{r}
#model <- vimure(random_net$X, random_net$R, mutuality=F, K=2, num_realisations=1, max_iter=150)
model <- vimure(X, random_net$R, mutuality=F, K=2, num_realisations=1, max_iter=150)
```


## Diagnostics

Use `summary` to produce result summaries of the results of the model. The output of `summary` is a R list that contains vimure's priors and posteriors and more. You can access the results to plot and analyse the final model. 

Open an issue in our [repository](https://github.com/latentnetworks/vimure/issues/new) to make request new metrics in `summary`.

```{r}
diag <- summary(model)
```

