---
title: "RMarkdown 4"
author: "Gabriela L. Borges"
date: "5/21/2022"
output: html_document
---

```{r setup, include=FALSE}
library(vimure)
library(igraph) 
library(ggplot2)
library(reshape2)
```

**OBJECTIVE:** To assess ViMuRe models on the 75 villages of Karnataka data.

**Change these values to run a different village+layer**
```{r}
village <- "vil55"
layer <- "visit"
```

# Data Input

Read villages as csv and convert to `vm.io.RealNetwork` or `graph` objects, convenient for our model usage.

```{r}
PATH <- "../../"
df_edges <- read.csv(paste0(PATH, "data/input/india_microfinance/formatted/", village, "_edges.csv"))

df_edges$type <- factor(df_edges$type, 
  levels=c(
    "borrowmoney","receivemoney", "receiveadvice", "helpdecision",
    "keroricego", "keroricecome", "visitgo", "visitcome"
    ), labels=c("money", "money", "advice", "advice", "kerorice", "kerorice", "visit", "visit"))

df_edges <- df_edges[df_edges$type %in% layer, ]
df_edges <- df_edges[df_edges$i != df_edges$j, ]
df_edges <- df_edges[!duplicated(df_edges), ]

head(df_edges)
```
 
## Using an igraph network

Vimure models expects as input a graph that can contains the _layer_, _reporter_ and _weight_ of each edge. 
Use `igraph::graph_from_data_frame` for create a graph using a dataset. The first two columns are the edgelist. Additional columns are considered as edge attributes (See more in ?graph_from_data_frame).

IMPORTANT: Vimure extracts these attributes by name, so it is crucial to rename the attributes before creating the graph (or parse the edgelist directly by `parse_graph_from_edgelist` and use params ego,alter,.. for mapping columns names - section below):
 
```{r}
edges <- df_edges[c(2:5, 1)] # Reorder columns
names(edges) <- c("Ego", "Alter", "layer", "weight", "reporter") # Rename
graph <- igraph::graph_from_data_frame(edges)

message(
  "Nodes: ", length(V(graph)),
  " | Edges: ", gsize(graph),
  " | Avg. degree: ", mean(degree(graph)), #directed graph,
  " | Reciprocity: ", reciprocity(graph)
)
```
Run the model:

```{r}
# Graph input
model <- vimure(graph, num_realisations=5L, max_iter=100L, seed=5)
```


 ## Parsing data to vimure classes
 
Vimure models expect as input the attributes `X` and `R` of a `parse_graph_from_edgelist` result. 
`parse_graph_from_edgelist` expects a dataframe with two required columns (_Ego_ and _Alter_) and three optional columns ( _layer_, _reporter_ and _weight_). Rename the dataframe's columns or use params ego,alter,... of `parse_graph_from_edgelist` for mapping column names to convert.

 
```{r}
vm_graph <- parse_graph_from_edgelist(df_edges, ego="i", alter="j", reporter="respondent")
```
 
 Run the model:
 
```{r}
# Vimure input
# model <- vimure(vm_graph$X, vm_graph$R, num_realisations=5L, max_iter=100L)
```
 
 
# Diagnostics

Use `summary` to produce result summaries of the results of the model. The output of `summary` is a R list that contains vimure's priors and posteriors and more. You can access the results to plot and analyse the final model. 

Open an issue in our [repository](https://github.com/latentnetworks/vimure/issues/new) to make request new metrics in `summary`.

```{r}
diag <- summary(model)
```
## Elbo values

```{r}
df_elbo <- diag$trace
ggplot(df_elbo, aes(x=iter, y=elbo, group=realisation, color=factor(realisation))) +
  geom_point() +
  geom_line(size=1.2) +
  theme_bw() +
  labs(x = "Iter", y="ELBO", color="Realisation")
```

## Estimated realibility

```{r}
plot_df <- melt(diag$reliability, id = "node")
ggplot(plot_df, aes(x = reorder(factor(node), -value), y=value)) +
  geom_point(alpha = 0.4, size=2, color = "#c45241") +
  theme_bw() +
  labs(x = "Nodes sorted from over-to under- reporting", y = "Realibility values") +
  theme(axis.text.x=element_blank())
```


