---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Vimure

<!-- badges: start -->
<!-- badges: end -->

Latent Network Models to Account for Noisy, Multiply-Reported Social Network Data in R.

## Installation

You can install the development version of vimure from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("latentnetworks/vimure", subdir="src/R", ref="18-vimure-v01-r-write-test_syntheticr")
```

## Setup (Development mode)

1. [Clone the repository](https://docs.github.com/en/repositories/creating-and-managing-repositories/cloning-a-repository) to a directory in your machine
2. Change to `vimure/src/R` directory
```bash
cd vimure/src/R
```
3. Open the Rstudio and load the package:
``` r
devtools::check()  # Build and check a source package
devtools::test()  # Run unittests
devtools::install()  # Install the current version of package
devtools::document()  # Generate man/ folder with functions docs
devtools::build_readme()  # Render the README.Rmd on a README.md file
```

## Usage Example

This is a basic example showing that virtualenv has been successfully configured 

```{r example}
library(vimure)
## basic example code

vimure:::vimureP  ## The Python package
```


