---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VIMuRe

<!-- badges: start -->
<!-- badges: end -->

Latent Network Models to Account for Noisy, Multiply-Reported Social Network Data.


If you use this code please cite this article (preprint).

> De Bacco C, Contisciani M, Cardoso-Silva J, Safdari H, Baptista D, Sweet T, Young JG, Koster J, Ross CT, McElreath R, Redhead D. Latent Network Models to Account for Noisy, Multiply-Reported Social Network Data. arXiv preprint [arXiv:2112.11396](https://arxiv.org/abs/2112.11396). 2021.

The VIMuRe R package wraps the [VIMuRe](https://github.com/latentnetworks/vimure/tree/18-vimure-v01-r-write-test_syntheticr/src/python) Python package. We use the [`reticulate`](https://rstudio.github.io/reticulate/) package to embeds a Python session within your R session, enabling seamless, high-performance interoperability. 

## Requirements

Vimure R package depends on R >= 3.3.0 and it is expect to work fine in all OS. 
The package also depends on Python >= 3.6, but you do not have to worry about that as we have a default set up that will run the first time you call `library(vimure)`. 

## Installation

You can install the development version of VIMuRe from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("latentnetworks/vimure", subdir="src/R", ref="develop")
```

### Default

If reticulate did not find a non-system installation of python you may be prompted if you want it to download and install miniconda. Miniconda is the recommended installation method for most users, as it ensures that the R python installation is isolated from other python installations. All python packages will by default be installed into a self-contained conda or venv environment named "r-vimure". Note that "conda" is the only supported method on Windows. 

If you initially declined the miniconda installation prompt, you can later manually install miniconda by running `reticulate::install_miniconda()`

### Set up your own Python Enviroment

If you do not want to install miniconda, you can set up your [Python enviroment](http://timsherratt.org/digital-heritage-handbook/docs/python-pip-virtualenv/) manually or by `reticulate::install_python()`. VIMuRe requires Python >= 3.6.

## Usage Example

Simply create an object with the desired synthetic network class:

```{r example}
library(vimure, quietly =T)
library(ggplot2, quietly =T)
library(ggcorrplot, quietly =T)
library(igraph, quietly =T)

random_net <- gm_Multitensor(N=50, M=50)
Y <- extract_Y(random_net) 
```

`vimure` is a R binding of a Python package. Many Python basic objects are quickly converted to R automatically.
Custom Python objects that can not be converted automatically are stored in R as a `python.builtin.object`.
As a `python.builtin.object`, you can access all object's attributes as it is in Python using the dollar sign `$`.

Use the function `class` to check if a object is stored in Python.

```{r}
class(random_net)
```

`random_net` is stored as a Python object. You can access its attributes using the dollar sign `$` or using our `extract_*` functions which always will return a R object.

```{r}
class(random_net$Y) # still a python object because it is a sptensor
class(extract_Y(random_net)) # extract_Y convert to array
```

Create a graph from the adjacency matrix and calculate some network statistics:

```{r}
graph <- graph_from_adjacency_matrix(Y[1, ,], mode = "directed")
paste0(
  "Nodes: ", length(V(graph)),
  " | Edges: ", gsize(graph),
  " | Avg. degree: ", mean(degree(graph)), # TODO: Change to directed graph
  " | Reciprocity: ", reciprocity(graph)
)
```

Given a network Y, we can generate N observed adjacency matrices as would have been reported by reporting nodes  ð‘š  ( ð‘šâˆˆð‘ ). This is achieved by the function `build_X`. Example:

```{r}
X <- build_X(random_net, flag_self_reporter=F, cutoff_X=F, mutuality=0.5, seed=20L)
Xavg <- extract_Xavg(random_net)

ggcorrplot(Xavg[1, ,]) + 
   scale_fill_gradient(low="white",high="#003396")
```



### Model

```{r}
model <- vimure(random_net$X, R=random_net$R, mutuality = T, num_realisations=1, max_iter=150)
diag <- summary(model, random_net)
```

## Setup (Development mode)

Use this setup if you want to modify anything in the package. For reproducibility reasons use the R version 4.1.2.

1. [Clone the repository](https://docs.github.com/en/repositories/creating-and-managing-repositories/cloning-a-repository) to a directory in your machine
2. In prompt, change to `vimure/src/R` directory and open a Rstudio session.
```bash
cd vimure/src/R
rstudio
```
3. In Rstudio, load the package:
``` r
devtools::install()  # Install the current version of package
```

